
---
title: "Data management for AIDA data"
author: "Kemmapon Chumchuen"
date: "2021-04-05"
output: html_document
---


## Load packages and dataset

- ข้อมูลตัวอย่างโหลดมาจาก [Google Form AIDA CRF](https://docs.google.com/forms/u/2/d/1p0AwxOOflxAQL8AvVncuBvGUMocvGtxQNRcl5_zJpAY/edit?usp=docs_home&ths=true)
```{r}
library(epicalc) 
library(readxl) 
library(plyr) 
library(splitstackshape) 


Sys.setlocale("LC_CTYPE", "thai")
options(encoding="UTF-8")
imp.template <- read.csv("ThailandAIDAData_ImportTemplate_2021-04-07.csv", header = T)
scr.dat <- read_excel("Siriraj-AIDA Screening (Sample Responses).xlsx", skip = 0, sheet = 1)
scs.dat <- read_excel("SCS datasheet3.xlsx", skip = 0, sheet = 1)

```

## Cleaning screening data

### Renaming variables of screening dataset

```{r}
str(scr.dat)
colnames(scr.dat)
imp.col.scr1 <- colnames(imp.template[2:23])
```
Replace the column names from google form data with ones from data import tool

```{r}
names(scr.dat) <- imp.col.scr1
colnames(scr.dat)
```

### Revalue the level of variables from screening dataset

```{r}
yn <- c("ใช่","ไม่ใช่")
hv <- c("เคย","ไม่เคย")

scr.dat[,2:22] <- as.data.frame(sapply(scr.dat[,2:22],
                                       mapvalues, from = yn, to = c(1,0)))
scr.dat[,2:22] <- as.data.frame(sapply(scr.dat[,2:22],
                                       mapvalues, from = hv, to = c(1,0)))
```

Check the modified values
```{r}
scr.dat[,1:3]
scr.dat[,6:12]
```
```{r}
scr.dat[,12:15] <- as.data.frame(ifelse(scr.dat[,12:15]=="ไม่ได้รับประทาน",0,1))
```

Check the modified value

```{r}
scr.dat[,11:15]
```

Deal with the "ไม่มี" in these 2 variables. Final format will depend on the decision made in the next meeting

```{r}
scr.dat[,16:17] <- as.data.frame(sapply(scr.dat[,16:17],
                                     mapvalues, from = "ไม่มี", to = NA))
scr.dat[,12:17]
```

Modified sex value
```{r}
scr.dat$sex_sc <- ifelse(scr.dat$sex_sc=="ชาย",1,2) 
colnames(scr.dat)
```

Screening Data is cleaned.

Back up data in rds format

```{r}
saveRDS(scr.dat,"cleaned scr.rds")
scr.dat <- readRDS("cleaned scr.rds")
```

Screening data is ready to import to REDCap template.


## Cleaning single cell sequencing data

### Renaming variables of sequencing data

```{r}
str(scs.dat)
colnames(scs.dat[1:21])
colnames(imp.template[25:53])
```
Replace the column names from google form data with ones from data import tool.

```{r}
imp.col.scr2 <- colnames(imp.template[25:53])

colnames(scs.dat)[1:29]  <- imp.col.scr2
colnames(scs.dat)
```

Generate columns for dx_disease to match REDCap template.

Changing variable name for concat.split.expanded function.
```{r}
names(scs.dat)[30] <- "dxnosis"
colnames(scs.dat)
```

Generating columns based on the level obtained
```{r}
add.diag.new <- scs.dat[,30] ; add.diag.new
add.diag.new <- concat.split.expanded(add.diag.new, "dxnosis", fill = 0, drop = TRUE,
                      sep = ",", type = "character")
colnames(add.diag.new)
```

These commands are for channing the name of obtained levels to match the template
```{r}
names(add.diag.new)[names(add.diag.new) == "dxnosis_ไม่เคยได้รับการวินิจฉัยว่าเป็นโรคข้างต้น"] <- "dx_disease___0"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคอัมพาตหรือโรคสมองขาดเลือด"] <- "dx_disease___1"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคไขมันในเลือดสูง (ไขมันในเลือดสูง/ ระดับคอเลสเทอรอล HDL ต่า)"] <- "dx_disease___2"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคความดันโลหิตสูง"] <- "dx_disease___3"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคเบาหวานหรือน้าตาลในเลือดสูง"] <- "dx_disease___4"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคหัวใจ (ลิ้นหัวใจรั่ว/ หัวใจล้มเหลว อื่น ๆ )"] <- "dx_disease___5"
names(add.diag.new)[names(add.diag.new) == "dxnosis_ภาวะหัวใจขาดเลือด"] <- "dx_disease___6"
names(add.diag.new)[names(add.diag.new) == "dxnosis_ภาวะหัวใจเต้นผิดจังหวะ"] <- "dx_disease___7"
names(add.diag.new)[names(add.diag.new) == "dxnosis_ภาวะกล้ามเนื้อหัวใจตาย"] <- "dx_disease___8"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคตับจากแอลกอฮอล์"] <- "dx_disease___9"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคแผลในกระเพาะอาหาร"] <- "dx_disease___10"
names(add.diag.new)[names(add.diag.new) == "dxnosis_นิ่วในถุงน้าดี"] <- "dx_disease___11"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคเกาต์"] <- "dx_disease___12"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคกระดูกพรุน"] <- "dx_disease___13"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคปริทันต์อักเสบ"] <- "dx_disease___14"
names(add.diag.new)[names(add.diag.new) == "dxnosis_โรคแพ้ภูมิตนเอง"] <- "dx_disease___15"
colnames(add.diag.new)
```

For ordering multiple choice diagnosis variable
```{r}
#add.diag.new <- add.diag.new[,c(5,16,8,9,11,15,3,4,2,10,13,1,7,6,12,14)]
```

```{r}
add.diag.new
des(scs.dat)
```

Delete the google format column 
```{r}
scs.dat <- scs.dat[,-30]
```

Changing the rest of column names
```{r}
imp.col.scr3 <- colnames(imp.template[72:78])
colnames(scs.dat)[30:36]  <- imp.col.scr3
colnames(scs.dat)
```

Generate columns for taking_hormones to match REDCap template.

Changing variable name for concat.split.expanded function.

```{r}
names(scs.dat)[37] <- "drug.horm"
colnames(scs.dat)
```


Generating columns based on the level obtained
```{r}
add.diag.new2 <- scs.dat[,37]
add.diag.new2 <- concat.split.expanded(add.diag.new2, "drug.horm", fill = 0, drop = TRUE,
                                      sep = ",", type = "character")
colnames(add.diag.new2)
```

These commands are for channing the name of obtained levels to match the template
```{r}
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ไม่ได้รับประทานยาใดๆ"] <- "taking_hormones___0"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_โกรทฮอร์โมน"] <- "taking_hormones___1"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ฮอร์โมนอินซูลิน/ ยาโรคเบาหวาน"] <- "taking_hormones___2"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ผลิตภัณฑ์ยาจากพลาสมา"] <- "taking_hormones___3"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_อื่น ๆ (ยาแผนจีนโบราณ/ ยาอายุรเวท)"] <- "taking_hormones___4"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยานอนหลับ"] <- "taking_hormones___5"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยาลดความดันโลหิต (ยารักษาความดัน/ Diovan/ Renibase)"] <- "taking_hormones___6"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยาสเตียรอยด์"] <- "taking_hormones___7"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยาแก้แพ้ (Allergic drugs/ Antihistamines)"] <- "taking_hormones___8"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยากดภูมิคุ้มกัน"] <- "taking_hormones___9"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยาปฏิชีวนะ"] <- "taking_hormones___10"
names(add.diag.new2)[names(add.diag.new2) == "drug.horm_ยาลดไขมันในเลือด (ยาลดคอเลสเทอรอล ชนิด LDL/ Mevalotin/ ลิปิโอดอล)"] <- "taking_hormones___11"
colnames(add.diag.new2)
```


### These commands are used in case of having other drugs in the dataset 

Add a blank other column in

```{r}
add.diag.new2$taking_hormones_oth  <- 0
```

Create a vector to exclude the registered drug, crete new dataset with only additional other drugs

```{r}
col <- c("taking_hormones___0","taking_hormones___1","taking_hormones___2",
         "taking_hormones___3","taking_hormones___4","taking_hormones___5",
         "taking_hormones___6","taking_hormones___7","taking_hormones___8",
         "taking_hormones___9","taking_hormones___10","taking_hormones___11")

add.diag.new3 <-   (add.diag.new2[, -which(names(add.diag.new2) %in% col)])

add.diag.new3
```

Fill drug names with column names

```{r}
add.diag.new3$taking_hormones_oth <- apply(add.diag.new3, 1,function(x) paste(names(add.diag.new3)[x > 0], collapse = "-"))
```

Delete the column name prefix, and create another other column (binary) to match REDCap
```{r}
oth <- add.diag.new3$taking_hormones_oth
oth <- gsub(".*_(.*)", "\\1", oth)
add.diag.new3$taking_hormones_oth <- oth


add.diag.new3$taking_hormones___12 <- apply(add.diag.new3, 1,
                                            function(x) ifelse(any(x == 1), 1, 0))
```

Delete unnecessary column

```{r}
add.diag.new3 <- add.diag.new3[,c(4,3)]
add.diag.new3

add.diag.new2 <- add.diag.new2[,-c(6,11,15)]
add.diag.new2

des(scs.dat)
```

Delete the google format column and merged the new columns, **always check the colum number**

```{r}

scs.dat <- scs.dat[,-37]
```

Add 3 sets of newly generated columns

```{r}
scs.dat <- cbind(scs.dat[1:29],add.diag.new,scs.dat[30:36],add.diag.new2,
                 add.diag.new3,scs.dat[37:38])
des(scs.dat)

```

Checking
```{r}
scs.dat[,50:60]
```

Change the last variable name, **always check for the last column**, as it can be varied due to added levels
```{r}
names(scs.dat)[59] <- "taking_hormone_period"
names(scs.dat)[60] <- "taking_vaccine"
des(scs.dat)
```

### Revalue the levels of variables from sequencing dataset
**Always consider the number of columns when revalue, as it can change in every entry**


```{r}
scs.dat$sex <- ifelse(scs.dat$sex=="ชาย",1,2)

hl <- c("รู้สึกอ่อนเพลีย/เหนื่อย ขณะนั่งพัก","ไม่มีอาการใดๆขณะพัก แต่เหนื่อยง่ายสำหรับกิจกรรมเล็กๆน้อยๆในบ้าน",
        "ไม่มีอาการขณะพัก แต่อาจเหนื่อยหอบถ้าต้องเดินขึ้นบรรได เดินซื้ออาหาร","เดินขึ้นบันไดหรือวิ่งระยะสั้นได้")    

scs.dat$health_level <- mapvalues(scs.dat$health_level,from = hl, to = c(1,2,3,4))

scs.dat$health_level

des(scs.dat)

yn <- c("ใช่","ไม่ใช่")
hv <- c("เคย","ไม่เคย")

scs.dat[,9:60] <- as.data.frame(sapply(scs.dat[,9:60],
                                mapvalues, from = yn, to = c(1,0)))
scs.dat[,9:60] <- as.data.frame(sapply(scs.dat[,9:60],
                                       mapvalues, from = hv, to = c(1,0)))

sm <- c("ไม่เคยสูบ","หยุดสูบแล้ว","สูบอยู่")

scs.dat$smoking <- mapvalues(scs.dat$smoking,from = sm, to = c(0,1,2))

alc <- c("ไม่เคยดื่มเลย","เลิกดื่มแล้ว",
         "น้อยกว่าเดือนละครั้ง (เช่น 2-3 เดือนครั้ง หรือ นานกว่านั้นต่อครั้ง)","ประมาณเดือนละครั้ง","ดื่มเกือบทุกสัปดาห์","ดื่มมากกว่า 1 ครั้งต่อสัปดาห์")

scs.dat$alcohol <- mapvalues(scs.dat$alcohol, from = alc, to = c(1,2,3,4,5,6))

head(scs.dat)
```

Single Cell Sequence Data is cleaned

SaveRDS as a back up

```{r}
saveRDS(scs.dat,"cleaned scs.rds")
scr.dat <- readRDS("cleaned scr.rds")
```

This file is ready to import to REDCap template

## Merging cleaned data to REDCap template and export
```{r}
dat.ready.imp <- as.data.frame(cbind(scr.dat,scs.dat))

imp.dat <- rbind.fill(imp.template,dat.ready.imp)
```

In case of data entry via google form, don't forget to add record_id in R and delete the last extra column 

```{r}
imp.dat$record_id <- c(1:5)
imp.dat <- imp.dat[,-96]

```

Manage date and time variable, as class could be altered during merging 

```{r}
imp.dat$timestamp <- scs.dat$timestamp
imp.dat$timestamp_sc <- scr.dat$timestamp_sc
imp.dat$date_recording <- scs.dat$date_recording
```


Export in CSV format and upload this file to REDCap
```{r}
write.csv(imp.dat,file = "Data Formated for REDCap Importation.csv", fileEncoding = "UTF-8",
          na="", row.names = F)
```
